---
title: "Plots"
author: "Cullan Mackrel"
date: "2026-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Necessary Libraries
``` {r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(devtools)
library(splitr)
library(openair)
library(withr)
library(readr)
library(lubridate)
library(scales)
```

## H2S Over Time. Scatter Plot. Monthly Average, 2020, AMS 4 & AMS 19

```{r warning=FALSE}
setwd("C:/Users/culla/Documents/data")
MonthPlot <- LongH2S %>%
mutate(LongH2S, date = as.Date(LongH2S$date)) %>%
timeAverage(avg.time = "month",
            type = "Station")

ggplot(MonthPlot %>% 
         selectByDate(year = 2020),
       aes(x = date, y = H2SConc, colour = Station)) + 
      geom_point() + 
  geom_line() +
  labs(x = "Months Of 2020", y = "Mixing Ratio (ppb)", title = "H2S Mixing Ratio (ppb) Throughout 2020") +
  scale_x_date(labels = date_format("%b"), breaks = "1 month") +
  theme_classic() 
```

## H2S Over Time. Marginal Plot. Monthly Average, 2020, AMS 4 & AMS 19, Marginal

```{r warning=FALSE}
setwd("C:/Users/culla/Documents/data")
MarginalPlotdata <- LongH2S %>%
  separate(col = date, into = c("date", "Hour"), sep = " ") %>%
  mutate(LongH2S, date = as.Date(LongH2S$date)) %>%
  timeAverage(avg.time = "month",
              type = "Station")
  
MarginalPlot <- ggplot(MarginalPlotdata %>%
         selectByDate(year = 2020),
       aes(x = date, y = H2SConc, colour = Station)) +
  labs(x = "Months", 
       y = "Mixing Ratio (ppb)",
       title = "Monthly Average H2S Mixing Ratio (ppb) in 2020") + 
  scale_x_date(labels = date_format("%b"), breaks = "1 month") +
  geom_point() +
  geom_line() +
  theme_classic() 
  

ggExtra::ggMarginal(MarginalPlot, margins = "y", groupColour = TRUE, groupFill = TRUE) 

```

## H2S Over Time. Scatter Plot. Monthly Average, 2020-2025, AMS 4 & Ams 19

```{r warning=FALSE}
setwd("C:/Users/culla/Documents/data")
Year1Plot <- LongH2S %>%
  separate(col = date, into = c("date", "Hour"), sep = " ") %>%
  mutate(LongH2S, date = as.Date(date)) %>%
  timeAverage(avg.time = "month",
              type = "Station")

ggplot(Year1Plot %>%
         selectByDate(year = 2020:2024),
       aes(x = date, y = H2SConc, colour = Station)) +
  labs(x = "Years", 
       y = "Mixing Ratio (ppb)",
       title = "Monthly Average H2S Mixing Ratio (ppb) From 2020-2025") + 
  scale_x_date(labels = date_format("%Y"), breaks = "1 year") +
  geom_point() +
  geom_line() +
  theme_classic()
```

## H2S Over Time. Scatter Plot. Yearly Average, 2020-2025, AMS 4 & 19

```{r warning=FALSE}
setwd("C:/Users/culla/Documents/data")
Year2Plot <- LongH2S %>%
  separate(col = date, into = c("date", "Hour"), sep = " ") %>%
  mutate(LongH2S, date = as.Date(date)) %>%
  timeAverage(avg.time = "year",
              type = "Station")

ggplot(Year2Plot %>%
         selectByDate(year = 2020:2024),
       aes(x = date, y = H2SConc, colour = Station)) +
  labs(x = "Years", 
       y = "Mixing Ratio (ppb)",
       title = "Yearly Average H2S Mixing Ratio (ppb) From 2020-2025") + 
  scale_x_date(labels = date_format("%Y"), breaks = "1 year") +
  geom_point() +
  geom_line() +
  theme_classic()
```

## H2S Over Time. Bubble Plot. 2020, Showing Fire Influence, AMS 4 & 19

```{r warning=FALSE}
 #Forming An H2S conc and fire influence table
setwd("C:/Users/culla/Documents/data")
 StationMerge <- StationH2S %>%
   separate(col = date, into = c("date", "hour"), sep = " ") %>%
   selectByDate(year = 2020) %>%
   unite(col = date,
         c("date", "hour"),
         sep = " ")
 setwd("E:/")
 FireData <- read_csv("2020AQData.csv") 
 AQMerge <- rename(FireData, hour = Hour) %>%
   mutate(hour = as.character(hour)) %>%
   unite(col = date,
         c("date", "hour"),
         sep = " ")

 MergedData <- left_join(StationMerge, AQMerge, by = "date") %>%
  pivot_longer(cols = c("AMS 4", "AMS 19"),
               names_to = "Station",
               values_to = "H2SConc")

#Bubble plot with fire influence 
BubblePlot <- MergedData %>%
  separate(col = date, into = c("date", "hour"), sep = " ") %>%
  mutate(date = as.Date(date)) 

 
 ggplot(BubblePlot, aes(x = date, y = H2SConc, size = Fire_Influence, colour = Station)) +
   geom_point(alpha = 0.5) +
   theme_classic() +
   labs(x = "2020", y = "H2S Mixing Ratio (ppb)", title = "Hydrogen Sulfide From Different Stations Throughout 2020", size = "Influenced By Wildfires?")
   
 
 print(BubblePlot)


```

## H2S Over Time. Scatter Plot. 2020, Showing Fire Influence AMS 4
```{r warning=FALSE}
 BubbleMerge <- StationH2S %>%
   mutate(date = paste(as.character(date)),
          date = as.POSIXct(date, format = "%Y-%m-%d %H")) %>%
   selectByDate(year = 2020) %>%
   select(date, `AMS 4`)
   
  FireMerge <- FireDataAMS4 %>%
    unite(col = date,
          c("date", "Hour"),
          sep = " ") %>% 
    mutate(date = paste(as.character(date)),
           date = as.POSIXct(date, format = "%Y-%m-%d %H")) %>%
    select(date, Fire_Influence)

  BubblePlot <- left_join(BubbleMerge, FireMerge, by = "date")  
  
  ggplot(BubblePlot, aes(x = date, y = `AMS 4`, colour = Fire_Influence)) + 
    geom_point() +
    labs(x = "2020", y = "H2S Mixing Ratio (ppb)", colour = "Fire Influence") +
    theme_classic()

```

## Number Of Influencing Fires Over Time. Scatter Plot. Monthly Average, 2020, AMS 4

```{r warning=FALSE}
setwd("E:/")
FiresWithInfluence <- FireData %>%
    select(date, Number_of_Fires, Fire_Influence) %>%
    filter(Fire_Influence == "Yes") %>%
    mutate(date = as.Date(date)) %>%
    timeAverage(avg.time = "month")

ggplot(FiresWithInfluence, aes(x = date, y = Number_of_Fires)) + 
  geom_point() +
  geom_line() +
  labs(x = "Months of 2020", y = "Number Of Fires With Influence", title = "A Monthly Average Of The Number Of Fires With Influence In 2020") +
  theme_classic()
```

## Intercept Time Of Fires Over Time. Scatter Plot. Monthly Average, 2020, AMS 4

``` {r warning=FALSE}
setwd("E:/")

InterceptPlotA <- FireData %>%
  select(c("date", "intercept_time")) %>%
  mutate(date = as.Date(date)) %>%
  timeAverage(avg.time = "month")

ggplot(InterceptPlotA, aes(x = date, y = intercept_time))  +
  geom_point() + 
  geom_line() +
  scale_y_reverse() +
  labs(x = "Months of 2020",y = "Intercept Time (hr)", title = "Average Time of Interception For The Months Of 2020 ") +
  theme_classic()
```

## Intercept Time Of Influencing Fires Over Time. Scatter Plot. Monthly Average, 2020, AMS 4

``` {r warning=FALSE}
setwd("E:/")
InterceptPlotB <- FireData %>%
  select(c("date", "Fire_Influence", "intercept_time")) %>%
  mutate(date = as.Date(date)) %>%
  filter(Fire_Influence == 'Yes') %>%
  timeAverage(avg.time = "month")

ggplot(InterceptPlotB, aes(x = date, y = intercept_time))  +
  geom_point() + 
  geom_line() +
  scale_y_reverse() +
  labs(x = "Months of 2020",y = "Intercept Time (hr)", title = "Average Intercept Time of Fires With Influence For The Months Of 2020 ") +
  theme_classic()
```

## H2S Influenced By Fires Over Time. Scatter Plot. Monthly Average, 2020, AMS 4

```{r warning=FALSE}
YesPlot <- MergedData %>%
  separate(col = date, into = c("date", "hour"), sep = " ") %>%
  mutate(date = as.Date(date)) %>%
  filter(Fire_Influence == 'Yes') %>%
  timeAverage(avg.time = "month")

ggplot(YesPlot, aes(x = date, y = H2SConc)) +
  geom_point() +
  labs(x = "2020", y = "H2S Mixing Ratio (ppb)", title = "H2S (ppb) Affected By Wildfires Throughout 2020") +
  geom_line() +
  theme_classic()

```

## H2S Not Influenced By Fires Over Time. Scatter Plot. Monthly Average, 2020, AMS 4
```{r warning=FALSE}
NoPlot <- MergedData %>%
  separate(col = date, into = c("date", "hour"), sep = " ") %>%
  mutate(date = as.Date(date)) %>%
  filter(Fire_Influence == 'No') %>%
  timeAverage(avg.time = "month")

ggplot(NoPlot, aes(x = date, y = H2SConc)) +
  geom_point() +
  labs(x = "2020", y = "H2S Mixing Ratio (ppb)", title = "H2S (ppb) Measurements Without Wildfire Influence Throughout 2020") +
  geom_line() +
  theme_classic()

```

## H2S Influenced By Fires Over H2S Not Influenced By Fires. Scatter Plot. Monthly Average, Coloured By Season, 2020, AMS 4

``` {r warning=FALSE}
H2SVsPlot <- MergedData %>%
  select(day, Fire_Influence, H2SConc) %>%
  rename(date = day) %>%
  mutate(date = as.Date(date)) %>%
  timeAverage(avg.time = "month",
              type = "Fire_Influence") %>%
  pivot_wider(names_from = Fire_Influence,
              values_from = H2SConc) %>%
  rename(Influenced = Yes,
         Not_Influenced = No) %>%
  mutate(date = as.character(date)) %>%
  mutate(season = case_when(date == "2020-01-01" ~ "Winter",
                            date == "2020-02-01" ~ "Winter",
                            date == "2020-03-01" ~ "Spring",
                            date == "2020-04-01" ~ "Spring",
                            date == "2020-05-01" ~ "Spring",
                            date == "2020-06-01" ~ "Summer",
                            date == "2020-07-01" ~ "Summer",
                            date == "2020-08-01" ~ "Summer",
                            date == "2020-09-01" ~ "Autumn",
                            date == "2020-10-01" ~ "Autumn",
                            date == "2020-11-01" ~ "Autumn",
                            date == "2020-12-01" ~ "Winter"))

ggplot(H2SVsPlot, aes(x = Not_Influenced, y = Influenced, colour = season, size = season)) +
  geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  theme_classic()
```

## H2S Distribution With And Without Fire Influence. Violin Plot. No Time Average, 2020, AMS 4, Outliers Removed

``` {r warning=FALSE}

Violin <- MergedData %>%
  select(date, Fire_Influence, H2SConc) %>%
  rename(Influence = Fire_Influence)
Q1 <- quantile(Violin$H2SConc, 0.25, na.rm = TRUE)
Q3 <- quantile(Violin$H2SConc, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
CleanViolin <- subset(Violin, Violin$H2SConc >= lower_bound & Violin$H2SConc <= upper_bound)

sample_size = CleanViolin %>% 
  group_by(Influence) %>% 
  summarize(num=n())

CleanViolin <- CleanViolin %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(Influence, "\n", "n=", num)) 


ggplot(na.omit(CleanViolin), aes(x = myaxis, y = H2SConc, fill = Influence)) +
  geom_violin() +
  labs(x = "Influenced By Wildfires?", y = "H2S Mixing Ratio (ppb)",
       title = "Distribution of Hydrogen Sulfide Measurements Depending On Fire Influence") +
  theme_classic()
  



```

### Adding Boxplots to the previous Violin plot

``` {r warning=FALSE}
Violin <- MergedData %>%
  select(Fire_Influence, H2SConc) %>%
  rename(Influence = Fire_Influence)
Q1 <- quantile(Violin$H2SConc, 0.25, na.rm = TRUE)
Q3 <- quantile(Violin$H2SConc, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
CleanViolin <- subset(Violin, Violin$H2SConc >= lower_bound & Violin$H2SConc <= upper_bound)

sample_size = CleanViolin %>% 
  group_by(Influence) %>% 
  summarize(num=n())

CleanViolin <- CleanViolin %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(Influence, "\n", "n=", num)) 

ggplot(na.omit(CleanViolin), aes(x = myaxis, y = H2SConc, fill = Influence)) +
  geom_violin() +
  geom_boxplot(width = 0.3, colour = "black", alpha = 0.2) +
  labs(x = "Influenced By Fires?", y = "H2S Mixing Ratio (ppb)", 
       title = "Distribution of Hydrogen Sulfide Measurements Depending On Fire Influence") +
  theme_classic()
  
```

## H2S Distribution With And Without Fire Influence. Violin Plot. 2020 Wildfire Season (May-Sep), AMS 4, Outliers Removed
```{r warning=FALSE}

Violin <- MergedData %>%
  select(date, Fire_Influence, H2SConc) %>%
  rename(Influence = Fire_Influence)
Q1 <- quantile(Violin$H2SConc, 0.25, na.rm = TRUE)
Q3 <- quantile(Violin$H2SConc, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
CleanViolin <- subset(Violin, Violin$H2SConc >= lower_bound & Violin$H2SConc <= upper_bound)

sample_size = CleanViolin %>% 
  group_by(Influence) %>% 
  summarize(num=n())

CleanViolin <- CleanViolin %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(Influence, "\n", "n=", num)) 



CleanViolin <- selectByDate(CleanViolin, month = 5:9)
  

ggplot(na.omit(CleanViolin), aes(x = Influence, y = H2SConc, fill = Influence)) +
  geom_violin() +
  labs(x = "Influenced By Wildfires?", y = "H2S Mixing Ratio (ppb)",
       title = "Distribution of Hydrogen Sulfide Measurements Depending On Fire Influence") +
  theme_classic()
```

## Pollutants Over
```{r warning=FALSE}
Pollutant <- AllData %>%
  select(date, H2SConc, PM2.5, SO2)

